### 2021-02-13 yfcc100m/paper/vldb2021/eval.tex
Add build plot and update exp setup

Given that there are no other open-source systems
that provides similar functionality and interfaces as VDMS,
we have implemented an equivalent visual data management system as a baseline,
comprised of a combination of widely available, off-the-shelf components:
MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
We decided to use a relational database instead of a non-relational
database because~\cite{Jatana2012, li_2019}:
\begin{itemize}
    \item Relational databases support atomicity, consistency, isolation,
    and durability (ACID) while non-relational may compromise some ACID properties.
    \item The YFCC100m data is clearly structured.
    \item We need to efficiently collate and return metadata records.
\end{itemize}
The baseline implementation only partially replicates the functionalities
that VDMS offers when it comes to image and metadata handling, built for the
purpose of an ad-hoc image search implementation.
This implementation is based upon internal tools used for
ML-based pipelines for media.
We have implemented a set of client-side applications that take care
of retrieving the components from the different systems, and applies
pre-processing operations when needed.
For visual data workloads, building an approach like the one
implemented as a baseline for this work
is the common practice in the industry \cite{haystack, tao}.

For all our experiments, we use two servers, one hosting a VDMS server and
another hosting the baseline implementation.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with hyper-threading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL has 256GB of DDR4 DRAM, while the server hosting VDMS
has 64GB of DDR4 DRAM.
We decided to run VDMS server in the machine with less DRAM to make
sure MySQL had no disadvantage, and because previous evaluation
indicated smaller footprint in the case of VDMS when
compared to similar baselines based on MySQL.
Other than the difference in DRAM space, machines are identical.
Both servers run Ubuntu 16.04.
The client application running the queries and measuring round-trip time
is connected to the server through a 1GB wired link through
a 10GB back-plane switch, same as both servers.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (retrieves metadata and
images) with VDMS (left) and the baseline (right).
The client application was implemented using Python 3 for both VDMS and
the baseline.

It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both
Apache WebServer and VDMS have direct access.
In the case of Videos, only the first frame is used for the image search.
More information and evaluation on video-specific functionalities can be found
in the appendix of this work.
In the case of the baseline, metadata is
stored in MySQL using an attached SSD disk.
Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of
comparison with respect to MySQL, which was not designed for
Persistent Memory type of storage.
The benefits of Persistent Memory for metadata and a full evaluation of
the PMGD subsystem is material for another paper,
and outside the scope of this evaluation.
For this experiment, in the case of VDMS we simply use a similar
attached SSD disk to store metadata.
Even if PMGD, the graph database used by VDMS, is designed for persistent memory,
it can deliver good performance when using SSDs directly, while still
providing ACID-compliant transactions.

For the metadata, we built VDMS and MySQL databases
using the YFCC100M dataset with incremental database sizes.
For simplicity, we named the database based on the approximate number of images
it contains, as follows: 1M, 5M, 10M, 50M, 100M.
The VDMS and MySQL databases have comparable number of elements.
The exact number of images/elements in each database are shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}.
The differences can be attributed to failures in data
preparation/loading because of incomplete/inconsistent formatting,
which is common in large datasets \cite{failures}.
In our set up, that difference is very small:
less than 0.1\% in terms of number of elements (images and/or metadata information).

@@

\textbf{MySQL/Baseline:}
Each MySQL database is created in a similar manner as VDMS
but the data is represented as three tables, following the relational model:
1) \textit{images} table: contains one row per image,
and a column for each property
associated with the images (some of which are listed in Section \ref{dataset});
2) \textit{taglist} table: contains one row per autotag element
(always 1,570 rows);
3) \textit{autotags} table: contains one row per autotag
assigned to an image. Each row contains a foreign key to the
image, a foreign key to the tag, and
the probability assigned to that tag belonging to that image.
Given that there are 8 autotags, on average, per image, the \textit{autotags}
table has around 8 times the number of rows present in the
\textit{metadata} table, as can be seen in Table~\ref{table:mysqltables}.
Using a Python client and simple queries, the \textit{taglist}
table is read from the list of tags with an auto-incremented
\textit{tagid} as a primary key, and the metadata table
is read from the YFCC100M metadata using the identifier as a primary key.
The \textit{autotags} table contains the generated autotags and
probabilities for entries of the \textit{images} table.
To generate the table, we split the \textit{autotags} data for each database
by the image identifier and autotag into new files.
The new files are read into the \textit{autotags} table with the image
identifier and \textit{tagid} as foreign keys.

@@

\begin{table}[ht]
\caption{MySQL Database - Number of Rows in each Table}
\centering
\begin{tabular}{c c c c}
\hline\hline
 & \multicolumn{3}{c}{Table}\\
\cline{2-4}
DB Name & images & autotags & taglist\\
\hline
% 100k & 100,000    & 848,912     & 1,570\\
% 500k & 498,707    & 4,241,200   & 1,570\\
1M   & 1,000,000  & 8,508,380   & 1,570\\
5M   & 4,987,379  & 42,425,905  & 1,570\\
10M  & 10,000,000 & 85,095,265  & 1,570\\
50M  & 50,000,000 & 425,446,208 & 1,570\\
100M & 99,206,564 & 896,002,496 & 1,570\\
\hline
\end{tabular}
\label{table:mysqltables}
\end{table}

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/db_time_size}
\caption{Time to build and size (in GB) of MySQL and VDMS databases.}
\label{fig:db_time_size}
\end{figure}

@@

Another important aspect to note is that
VDMS requires more storage for metadata, shown in Figure~\ref{fig:db_time_size}.
This is space used to store information about each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
especially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 30-41\% more storage than MySQL for storing the same amount
of metadata.
This may become a factor if storage is a limitation, but it should also be noted
that even if we have a 41\% increase in metadata size,
metadata accounts for less than 2\% of the overall database size.
For example, the largest database (both metadata and images) we built
(100M) has around 230GB of metadata and 12TB of images.
In systems where persistent memory is a scarce resource,
the increased storage foot print of PMGD may represent a challenge.
On the other hand, persistent memory is expected to be available
in the order of TBs per server, which should fit the
metadata of intensive use-cases\cite{IntelXPoint15}.

@@

In order to evaluate VDMS and the baseline on our use-case queries,
we implemented 5 queries that filter and retrieve a specific set of images.
We chose these queries because they represent typical use-cases where a
cohort of images is to be retrieved and processed from a large corpus of data.
As we mentioned before, we took this approach due to the lack of standard
benchmarks that are oriented towards visual data retrieval.
We use the metadata associated with the images to filter said images.

@@

It is important to note that when querying for images with certain
\textit{autotags}, we also apply a filter using the probability.
For instance, we only retrieve images with an autotag \textit{alligator}
and a probability higher than 92\%.
These probabilities are both present in VDMS (in the form of a property
of the \textit{connection} between the image and that \textit{autotag}),
as well as in MySQL (in the form of a column in the \textit{autotags} table
that links images with tags).
In the case of VDMS, the query involves a graph traversal query that starts
from the \textit{autotag} node and ends in the images node,
following \textit{connections} between the image and that \textit{autotag}).
In the case of the baseline implementation,
the query involves JOIN operations between the 3 tables.
The implementation of this evaluation, as well as all the queries, are available
under the benchmarks branch of the VDMS project
\footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
and some examples of the queries can be found in the appendix of this
paper.

@@

Also, note that the size of the result (number of images retrieved)
is linear with the size of the database. This is, if a query returns 100 images
for the 1M database, it will return around 1000 images for the 10M database.
This poses a problem when evaluating performance as the size of the database increase,
and clearly understanding the measurements.
Because of this reason, we control the number of returned images for all the
databases using the probability of the \textit{autotags}
(higher probabilities returns less images), so that the queries in
this experiment return a similar number of images for all database sizes.
In other words, as the size of the database increase, we increase the probability
threshold for the queries. We do this for both VDMS and the baseline, of course.
This way, we remove bottleneck introduced by network bandwidth that would
otherwise over-complicate the understanding of the results.

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{figures/plot_56_mosaic_query_times}
\caption{Performance Analysis using all queries from our use-case
described in the Experimental Setup Section.
We show queries in different figures for readability reasons.
Top figures show the throughput of the queries for
just retrieving metadata associated with the images.
Bottom figures show the throughput when retrieving both metadata and images,
plus operations applied to images when applicable (queries 2, 3, 4 and 5).
The experiments show the performance of both systems (VDMS and baseline) as the
database size increases.
These queries were run using 56 simultaneous clients (nthr = 56),
and averaged out of 10 runs.}
\label{fig:q_performance_56}
\end{figure*}

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{figures/plot_56_mosaic_results_throughput}
\caption{Throughput Analysis using all queries from our use-case
described in the Experimental Setup Section.
We show queries in different figures for readability reasons.
Top figures show the throughput of the queries for
just retrieving metadata associated with the images.
Bottom figures show the throughput when retrieving both metadata and images,
plus operations applied to images when applicable (queries 2, 3, 4 and 5).
The experiments show the performance of both systems (VDMS and baseline) as the
database size increases.
These queries were run using 56 simultaneous clients (nthr = 56),
and averaged out of 10 runs.}
\label{fig:q_throughput_56}
\end{figure*}

@@

\subsubsection{Concurrency Analysis}

>>>>

Given that there are no other open-source systems
that provides similar functionality and interfaces as VDMS,
we have implemented two equivalent visual data management system as baselines,
comprised of a combination of widely available, off-the-shelf components.
The first baseline uses
MySQL Server 5.7 (for storing metadata),
Apache Web Server 2.4.18 (as interface for image access), and
OpenCV 3.3 (to provide pre-processing operations on images).
The other baseline system replaces the MySQL Server 5.7 with the
most advanced open source relational database, PostgreSQL 9.5 \cite{postgresql}.
We decided to use relational databases in the baseline systems instead of non-relational
databases because~\cite{Jatana2012, li_2019}:
\begin{itemize}
    \item Relational databases support atomicity, consistency, isolation,
    and durability (ACID) while non-relational may compromise some ACID properties.
    \item The YFCC100m data is clearly structured.
    \item We need to efficiently collate and return metadata records.
\end{itemize}
The baseline implementations only partially replicate the functionalities
that VDMS offers when it comes to image and metadata handling, built for the
purpose of an ad-hoc image search implementation.
This implementation is based upon internal tools used for
ML-based pipelines for media.
We have implemented a set of client-side applications that take care
of retrieving the components from the different systems, and applies
pre-processing operations when needed.
For visual data workloads, building a system like the baselines for this work
is the common practice in the industry \cite{haystack, tao}.

For all our experiments, we use two servers with Ubuntu 16.04, one hosting a VDMS server and
another hosting the baseline implementations.
Both servers have a dual-socket Intel\textsuperscript{\textregistered}
Xeon\textsuperscript{\textregistered} Platinum 8180 CPU @ 2.50GHz (Skylake),
each CPU with 28 physical cores with hyper-threading enabled,
for a total of 112 logical cores per server.
The server hosting MySQL and PostgreSQL has 256GB of DDR4 DRAM, while the server hosting VDMS
has 64GB of DDR4 DRAM.
We decided to run VDMS server in the machine with less DRAM to make
sure MySQL and PostgreSQL had no disadvantage, and because previous evaluation
indicated smaller footprint in the case of VDMS when
compared to similar baselines based on MySQL.
To build both MySQL and PostgreSQL on the same machine,
we stored each relational database on separate SSD drives.
Other than the difference in DRAM space and storage needed for two baselines, the machines are identical.
The client application running the queries and measuring round-trip time
is connected to the server through a 1GB wired link through
a 10GB back-plane switch, same as both servers.
The client application was implemented using Python 3 for both VDMS and the baselines.
Figure~\ref{fig:systems} shows a logical view of the difference between the
interaction of the client application (retrieves metadata and
images) with VDMS (left) and the MySQL baseline (right).

It is worth noting that the images are stored in a shared repository
(ext4 filesystem on a RAID 6 configuration of 16TB) that both
Apache WebServer and VDMS have direct access.
In the case of Videos, only the first frame is used for the image search.
More information and evaluation on video-specific functionalities can be found
in the appendix of this work.
In the case of the baselines, metadata is
stored in MySQL and PostgreSQL using an attached SSD disk.
Even if VDMS has native support for Optane Persistent Memory,
we do not use it in this experiment because of fairness of
comparison with respect to MySQL and PostgreSQL, which were not designed for
Persistent Memory type of storage.
The benefits of Persistent Memory for metadata and a full evaluation of
the PMGD subsystem is material for another paper,
and outside the scope of this evaluation.
For this experiment, in the case of VDMS we simply use a similar
attached SSD disk to store metadata.
Even if PMGD, the graph database used by VDMS, is designed for persistent memory,
it can deliver good performance when using SSDs directly, while still
providing ACID-compliant transactions.

For the metadata, we built VDMS, MySQL, and PostgreSQL databases
using the YFCC100M dataset with incremental database sizes.
For simplicity, we named the database based on the approximate number of images
it contains, as follows: 1M, 5M, 10M, 50M, 100M.
The databases have comparable number of elements.
The exact number of images/elements in each database are shown in
Table~\ref{table:vdmsnodes} and ~\ref{table:mysqltables}.
The differences can be attributed to failures in data
preparation/loading because of incomplete/inconsistent formatting,
which is common in large datasets \cite{failures}.
In our set up, that difference is very small:
less than 0.253\% in terms of number of elements (images and/or metadata information).

@@

\textbf{MySQL/Baseline 1:}
Each MySQL database is created in a similar manner as VDMS
but the data is represented as three tables, following the relational model:
1) \textit{images} table: contains one row per image,
and a column for each property
associated with the images (some of which are listed in Section \ref{dataset});
2) \textit{taglist} table: contains one row per autotag element
(always 1,570 rows);
3) \textit{autotags} table: contains one row per autotag
assigned to an image. Each row contains a foreign key to the
image, a foreign key to the tag, and
the probability assigned to that tag belonging to that image.
Given that there are 8 autotags, on average, per image, the \textit{autotags}
table has around 8 times the number of rows present in the
\textit{metadata} table, as can be seen in Table~\ref{table:mysqltables}.
Using a Python client and simple queries, the \textit{taglist}
table is read from the list of tags with an auto-incremented
\textit{tagid} as a primary key, and the metadata table
is read from the YFCC100M metadata using the identifier as a primary key.
The \textit{autotags} table contains the generated autotags and
probabilities for entries of the \textit{images} table.
To generate the table, we split the \textit{autotags} data for each database
by the image identifier and autotag into new files.
The new files are read into the \textit{autotags} table with the image
identifier and \textit{tagid} as foreign keys.

@@

\textbf{PostgreSQL/Baseline 2:}
Each PostgreSQL database is created in the exact same manner as MySQL
where data is represented as three tables:
\textit{images}, \textit{taglist}, and \textit{taglist}.
PostgreSQL works well out-of-box and optimizations were not necessary,
as in mySQL, to load or execute queries.
When executing queries for larger databases,
it is pertinent to have enough disk space for both
the database and tablespace.
The tablespace is associated with a database and stores the temporary files
created within the database object ~\cite{postgresql}.
When processing the larger databases, additional
storage may be needed for the tablespace or a diskfull error may occur. To avoid this,
we allocated a separate SSD disk specifically for the tablespace of all databases.

By default, PostgreSQL can manage concurrent access to data well using Multiversion
Concurrency Control (MVCC) and a Serializable Snapshot Isolation (SSI) level of transcation isolation.
MVCC provides better performance by using SSI to guarantee querying data never
blocks writing data, and vice versa ~\cite{postgresql}.
The PostgreSQL configuration file initially limits the maximum
number of concurrent connections to the database server to 100.
We increased this limit to 200 to complete the concurency analysis in \label{concurrency analysis}.

\begin{table}[ht]
\caption{MySQL and PostgreSQL Databases - Number of Rows in each Table}
\centering
\begin{tabular}{c c c c}
\hline\hline
 & \multicolumn{3}{c}{Table}\\
\cline{2-4}
DB Name & images & autotags & taglist\\
\hline
% 100k & 100,000    & 848,912     & 1,570\\
% 500k & 498,707    & 4,241,200   & 1,570\\
1M   & 1,000,000  & 8,508,380   & 1,570\\
5M   & 4,987,379  & 42,425,905  & 1,570\\
10M  & 10,000,000 & 85,095,265  & 1,570\\
50M  & 50,000,000 & 425,446,208 & 1,570\\
100M & 99,206,564 & 896,002,496 & 1,570\\
\hline
\end{tabular}
\label{table:mysqltables}
\end{table}

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{figures/all_build_times_plot}
\caption{Time to build and size (in GB) of MySQL, PostgreSQL, and VDMS databases.}
\label{fig:db_time_size}
\end{figure}

@@

Another important aspect to note is the storage footprint of the databases.
PostgreSQL and VDMS require the same amount of storage for metadata which
is more than required by mySQL, shown in Figure~\ref{fig:db_time_size}. In the case of VDMS,
this is space used to store information about each node/connection.
The Graph Database internal to VDMS (called PMGD) was designed for performance,
especially in environments where persistent memory is present.
This design decision comes as a trade-off for storage footprint, which is
noticeable in our results.
VDMS required 30-41\% more storage than MySQL for storing the same amount
of metadata.
This may become a factor if storage is a limitation, but it should also be noted
that metadata accounts for less than 2\% of the overall database size
even if we have a 41\% increase in metadata size.
For example, the largest database (both metadata and images) we built
(100M) has around 230GB of metadata and 12TB of images.
In systems where persistent memory is a scarce resource,
the increased storage foot print of PMGD may represent a challenge.
On the other hand, persistent memory is expected to be available
in the order of TBs per server, which should fit the
metadata of intensive use-cases\cite{IntelXPoint15}.

@@

In order to evaluate VDMS and the baselines on our use-case queries,
we implemented 5 queries that filter and retrieve a specific set of images.
We chose these queries because they represent typical use-cases where a
cohort of images is to be retrieved and processed from a large corpus of data.
As we mentioned before, we took this approach due to the lack of standard
benchmarks that are oriented towards visual data retrieval.
We use the metadata associated with the images to filter said images.

@@

It is important to note that when querying for images with certain
\textit{autotags}, we also apply a filter using the probability.
For instance, we only retrieve images with an autotag \textit{alligator}
and a probability higher than 92\%.
These probabilities are both present in VDMS (in the form of a property
of the \textit{connection} between the image and that \textit{autotag}),
as well as in MySQL and PostgreSQL (in the form of a column in the \textit{autotags} table
that links images with tags).
In the case of VDMS, the query involves a graph traversal query that starts
from the \textit{autotag} node and ends in the images node,
following \textit{connections} between the image and that \textit{autotag}).
In the case of the baseline implementations,
the query involves JOIN operations between the 3 tables.
The implementation of this evaluation, as well as all the queries, are available
under the benchmarks branch of the VDMS project
\footnote{https://github.com/IntelLabs/vdms/ under benchmark/benckmarks/visual\_storm/yfcc100m},
and some examples of the queries can be found in the appendix of this
paper.

@@

Also, note that the size of the result (number of images retrieved)
is linear with the size of the database. This is, if a query returns 100 images
for the 1M database, it will return around 1000 images for the 10M database.
This poses a problem when evaluating performance as the size of the database increase,
and clearly understanding the measurements.
Because of this reason, we control the number of returned images for all the
databases using the probability of the \textit{autotags}
(higher probabilities returns less images), so that the queries in
this experiment return a similar number of images for all database sizes.
In other words, as the size of the database increase, we increase the probability
threshold for the queries. We do this for both VDMS and the baselines, of course.
This way, we remove bottleneck introduced by network bandwidth that would
otherwise over-complicate the understanding of the results.

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{figures/plot_56_mosaic_query_times}
\caption{Performance Analysis using all queries from our use-case
described in the Experimental Setup Section.
We show queries in different figures for readability reasons.
Top figures show the throughput of the queries for
just retrieving metadata associated with the images.
Bottom figures show the throughput when retrieving both metadata and images,
plus operations applied to images when applicable (queries 2, 3, 4 and 5).
The experiments show the performance of all systems (VDMS and two baselines) as the
database size increases.
These queries were run using 56 simultaneous clients (nthr = 56),
and averaged out of 10 runs.}
\label{fig:q_performance_56}
\end{figure*}

\begin{figure*}[ht]
\includegraphics[width=\textwidth]{figures/plot_56_mosaic_results_throughput}
\caption{Throughput Analysis using all queries from our use-case
described in the Experimental Setup Section.
We show queries in different figures for readability reasons.
Top figures show the throughput of the queries for
just retrieving metadata associated with the images.
Bottom figures show the throughput when retrieving both metadata and images,
plus operations applied to images when applicable (queries 2, 3, 4 and 5).
The experiments show the performance of all systems (VDMS and the two baselines) as the
database size increases.
These queries were run using 56 simultaneous clients (nthr = 56),
and averaged out of 10 runs.}
\label{fig:q_throughput_56}
\end{figure*}

@@

\subsubsection{Concurrency Analysis} \label{concurrency analysis}
